SMT-COMP 20243\\ 19th International Satisfiability Modulo Theory Competition

### 2024-07-18 2024/slides/presresults/slides.tex
SMT Workshop 2024 slides draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   SMT-COMP 2023 - 
%
%   1 hour
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%% Titel, Autor und Datum des Vortrags:
\title{SMT-COMP 2023\\
18th International Satisfiability Modulo Theory Competition}
\author{Martin Bromberger \and Jochen Hoenicke \and \emph{Fran\c{c}ois Bobot}}
\date{July 5, 2023}

%% Institut
\institute{
CEA List, France \and
MPI für Informatik, Germany \and
Certora, Israel
}

@@

  \begin{minipage}[b]{.6\textwidth}
    Annual competition for \emph{SMT solvers}\\
    on (a selection of) benchmarks from \emph{SMT-LIB}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \begin{block}{History}
      \begin{tabular}{rp{3cm}}
        \textbf{2005} & first competition \\
        \textbf{2013} & evaluation instead \\
        & of competition\\
        \textbf{2014} & since then hosted\\
        & by \emph{StarExec}
      \end{tabular}
    \end{block}
  \end{minipage}

  Goals:
  \begin{itemize}
  \item spur development of SMT solver implementations
  \item promote SMT solvers and their usage
  \item support the SMT-LIB project
    \begin{itemize}
    \item to promote and develop the SMT-LIB format
    \begin{itemize}
      \item model validation
      \item proof checking
    \end{itemize}
    \item to collect relevant benchmarks
    \end{itemize}
  \item engage and include new members
  \end{itemize}

@@

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{block}{Non-incremental}
        434\,212 instances {\small (+42849)}\\ % 434212-391363=42849
        with 1 query each \\
        in 83 logics {\small (+2)}.
      \end{block}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{block}{Incremental}
        43\,287 instances {\small (+2)}\\
        with 34\,036\,491 queries {\small (+37\,556)} \\ %34036491-33998935=37556
        in 38 logics. % -1 ???
      \end{block}
    \end{column}
    \begin{column}{0.08\textwidth}
    \end{column}
  \end{columns}
  \pause
    \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{block}{Selected Non-incremental}
        227\,940 instances
      \end{block}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{block}{Selected Incremental}
        22\,302 instances
      \end{block}
    \end{column}
    \begin{column}{0.08\textwidth}
    \end{column}
  \end{columns}

@@

  \emph{Cloud and Parallel Track} (sponsored by AWS, led by Mike Whalen)
  \begin{itemize}
  \item Solve a large problem over the cloud (or a big computer)
  \begin{itemize}
    \item 100 machines, 1600 cores, 6400 GB of memory (cloud)
    \item 64 cores, 256 GB of memory (parallel)
  \end{itemize}
  \item Solver answers sat/unsat/unknown
  \end{itemize}

  \pause\bigskip

  \emph{Proof Exhibition Track}
  \begin{itemize}
  \item Solver submitted together with a checker for unsatisfiability proofs
  \item No predefined format or checker
  \item No ranking
  \item Qualitative assessment
\end{itemize}

@@

\begin{frame}{Tracks, Solvers, Divisions, and Benchmarks}
  Teams: 25 (+4) %2022 21
  \bigskip

  \begin{tabular}{c|r@{}l|r@{}l|c}
    Track & \multicolumn{2}{c|}{Solvers} & \multicolumn{2}{c|}{Divisions}  & Benchmarks \\
    \hline
    Single Query  &  22&(=)  & 19&(=)  & 113\,139 \\
    Incremental &  7&(-1)   & 17&(=)  & 22\,301   \\
    Unsat Core  &  6&(=)   & 16&(-1)  & 72\,958  \\
    Model Validation  &  11&(+3)    &  13& (+ 6 exp.)  & 61\,083  \\
    Proof Exhibition  &  4&(=)    &  & 19 exp.  & 59\,114  \\
    \hline
    Parallel &   3&(-1)      &   &4 exp.  & 400 \\
    Cloud & 2&(-2)      &  &4 exp.  & 400 \\

@@

  Number in parenthesis shows changes from 2022
\end{frame}

@@

  SMT-COMP 2022 participants rely on multiple reasoning frameworks:
  \begin{itemize}
  \item CDCL(T), Saturation, MCSAT, CP
  \item automata
  \item finite domain
  \item local search
  \item besides wrappers extending the scope of existing solvers
  \end{itemize}

  \bigskip
  Seven new solvers participated:
  \begin{itemize}
\item iProver {\footnotesize (Korovin et al.)}
\item SMT-RAT-MCSAT {\footnotesize (Jasper Nalbach et al.)}
\item UltimateIntBlastingWrapper+SMTInterpol {\footnotesize (Max Barth et al.)}
\item Yaga {\footnotesize (Hanák et al.)}
\item Z3-alpha {\footnotesize (Lu et al.)}
\item Z3-Noodler {\footnotesize (Havlena et al.)}
\item Z3-Owl {\footnotesize (Ma et al.)}
\end{itemize}

@@

\myincludepdf{slides-bitwuzla.pdf}

\myincludepdf{slides-colibri.pdf}

\myincludepdf{slides-cvc5.pdf}

\myvideopdf{cvc5-NRA-LS}{slides-cvc5-nra-ls.mp4}

\myvideopdf{iProver}{slides-iProver.mp4}

\myvideopdf{ismt, Yices-ismt}{slides-ismt.mp4}

\myincludepdf{slides-ismt-ppt.pdf}

\myincludepdf{slides-opensmt.pdf}

\myvideopdf{Ostrich}{slides-Ostrich.mkv}

\myincludepdf{slides-smtinterpol.pdf}

\myincludepdf{slides-smtrat.pdf}

\myvideopdf{UltimateEliminator+MathSAT}{slides-UltimateEliminator.mkv}

\myincludepdf{slides-vampire.pdf}

\myincludepdf{slides-yaga.pdf}

\myincludepdf{slides-Yices-2023.pdf}

\myincludepdf{slides-YicesQS-2023.pdf}

%\myvideopdf{Z3-Z3++}{slides-z3-z3++.mp4}
\begin{frame}
  \frametitle{Z3-Z3++}

  \begin{center}
    \url{https://youtu.be/fBB0Wxxf9vA}
  \end{center}
\end{frame}

  \begin{frame}
    \frametitle{Other participants}
    \begin{itemize}
\item Q3B
\vitem Q3B-pBDD
\vitem STP
\vitem UltimateIntBlastingWrapper+SMTInterpol
\vitem z3-alpha
\vitem Z3-Noodler
\vitem Z3-Owl
    \end{itemize}
  \end{frame}

@@

  Submitted by organisers
  \begin{itemize}
  \item Best solvers, per division, from previous years (27 Solvers)
  \end{itemize}
  \bigskip

  Submitted by participants
  \begin{itemize}
  \item Fixed solvers (OSTRICH, Z3-Owl, Bitwuzla, Yices2, Z3-Noodler, iProver)  
  \end{itemize}
\end{frame}

@@

  \begin{itemize}
  \item  50411 benchmarks of 227\,938 have no status

  \vitem 300 benchmarks with disagreements (ABV,
  BV,
  LIA,
  NIA,
  QF\_ABV,
  QF\_AUFBV,
  QF\_BVFP,
  QF\_FP,
  QF\_NRA,
  QF\_SLIA,
  UF,
  UFDT,
  UFDTLIRA)

  \vitem We manually resolved the disagreements: authors confirmed solver
  unsoundness
  \vitem We had 10 solvers with soundness issues:
  \begin{columns}
    \column{0.45\textwidth}
    \begin{itemize}
      \item Bitwuzla (uc printing error)
      \item iProver (\& Fixed)
      \item OSTRICH
      \item Q3B
      \item UltimateEliminator+MathSAT
    \end{itemize} 
    \column{0.55\textwidth}
    \begin{itemize}
      \item UltimateIntBlastingWrapper+SMTInterpol
      \item Vampire
      \item Yices2 (\& Fixed)
      \item Z3-Noodler (\& Fixed)
      \item Z3-Owl (\& Fixed)
    \end{itemize} 
\end{columns}

  \end{itemize}
  \end{frame}

\begin{frame}[fragile]
  \frametitle{Plans for SMT-COMP 2023}
  \begin{itemize}
    \item "Revelations are found in clouds” --- Serge King
    \item MV: Algebraic number $\implies$ non experimental
\begin{verbatim}
(root-of-with-ordering (coeffs p_0 p_1 ... p_n) i)

@@

\begin{frame}{Plans for SMT-COMP 2023?}
  \begin{itemize}
    \item Proof \textit{validation} track

      \vitem Hopefully proof exhibition this year will help

      \vitem We have to analyze the data still
      \begin{itemize}
        \item Job only finished last week (took 18 days to run)
        \item 830gb
      \end{itemize}

@@

\begin{frame}{SMT-COMP organizing committee}
  Three people organize the SMT-COMP.  In 2023:
  \begin{itemize}
  \item Martin Bromberger
  \item Jochen Hoenicke
  \item Fran\c{c}ois Bobot
  \end{itemize}

  Jochen have been organizer for four-years! He can rest happy.
  \bigskip

  We need a successor for next year's competition.
  Contact us if you would like to volunteer!
\end{frame}

\begin{frame}{Acknowledgements}
  \begin{itemize}
    \item \emph{Andrea Micheli}: pysmt
    \item \emph{Guillaume Bury, FB.}: Model Validator
    \item \emph{Clark Barrett, Pascal Fontaine, Aina Niemetz, Mathias Preiner, Hans-Jörg Schurr}: SMT-LIB benchmarks
    \item \emph{Aaron Stump}: StarExec support
    \item \emph{Mike Whalen and team}: Cloud/Parallel Track
  \end{itemize}
  \bigskip

@@

\begin{frame}[shrink=0.95]{Benchmark contributors}
  In 2023 \emph{new benchmarks} were contributed by:

  \small
  \begin{itemize}
    \item  Alex Coffin
    \item  Alex Ozdemir
    \item  Ali Uncu, James Davenport and Matthew England
    \item  Bohan Li
    \item  Elizabeth Polgreen
    \item  Fuqi Jia
    \item  Johann-Tobias Aaron and Raphael Schäg
    \item  Matthew England and Miguel Del Rio Almajano
    \item  Nicolas Amat
    \item  Yannick Moy
    \item  Yoni Zohar
  \end{itemize}
\end{frame}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   SMT-COMP 2024 - 
%
%   1 hour
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%% Titel, Autor und Datum des Vortrags:
\title{SMT-COMP 20243\\
19th International Satisfiability Modulo Theory Competition}
\author{Fran\c{c}ois Bobot \and \emph{Martin Bromberger} \and Martin Jon\'{a}\v{s} }
\date{July 22, 2024}

%% Institut
\institute{
CEA List, France \and
MPI für Informatik, Germany \and
Masaryk University, Czechia
}

@@

  \begin{columns}
    \begin{column}{0.6\textwidth}
    Annual competition for \emph{SMT solvers}\\
    on (a selection of) benchmarks from \emph{SMT-LIB}

    \bigskip
    \bigskip

    Goals:
    \begin{itemize}
      \item spur development of SMT solver implementations
      \item promote SMT solvers and their usage
      \item support the SMT-LIB project
      \begin{itemize}
        \item to promote and develop the SMT-LIB format
        \begin{itemize}
          \item model validation
          \item proof checking
        \end{itemize}
        \item to collect relevant benchmarks
      \end{itemize}
      \item engage and include new members
    \end{itemize}
    \end{column}
    \begin{column}{0.4\textwidth}
    \begin{block}{History}
      \begin{tabular}{rp{3cm}}
        \textbf{2005} & first competition \\
        \textbf{2013} & evaluation instead \\
        & of competition\\
        \textbf{2014} & since then hosted\\
        & by \emph{StarExec}\\
        \textbf{2024} & since then hosted\\
        & by \emph{BenchExec} and the \emph{SoSy-Lab} 
      \end{tabular}
    \end{block}
    \end{column}
  \end{columns}

@@

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{block}{Non-incremental}
        438\,631 instances {\small (+4\,419)}\\ % 438631-434212=4419
        with 1 query each \\
        in 84 logics {\small (+1)}.
      \end{block}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{block}{Incremental}
        44\,333 instances {\small (+1\,046)}\\ % 44333-43287=1046
        with 34\,069\,145 queries {\small (+32\,654)} \\ %34036491-33998935=37556
        in 39 logics. % -1 ???
      \end{block}
    \end{column}
    \begin{column}{0.08\textwidth}
    \end{column}
  \end{columns}
  \pause
    \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{block}{Selected Non-incremental}
        249\,301 instances % old 227\,940???
      \end{block}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{block}{Selected Incremental}
        22\,826 instances % old 22\,302
      \end{block}
    \end{column}
    \begin{column}{0.08\textwidth}
    \end{column}
  \end{columns}

@@

  \emph{Cloud and Parallel Track} (sponsored by AWS, led by Robert Jones and Benjamin F Jones)
  \begin{itemize}
  \item Solve a large problem over the cloud (or a big computer)
  \begin{itemize}
    \item 100 machines, 1600 cores, 6400 GB of memory (cloud)
    \item 64 cores, 256 GB of memory (parallel)
  \end{itemize}
  \item Solver answers sat/unsat/unknown
  \item Thanks to Mike Whalen for managing these tracks the last two years and 
        thanks to Robert Jones and Benjamin F Jones for taking over
  \end{itemize}

@@

\begin{frame}[fragile]{SMT-COMP Tracks (discontinued)}

  \emph{Proof Exhibition Track}
  \begin{itemize}
  \item Solver submitted together with a checker for unsatisfiability proofs
  \item No predefined format or checker
  \item No ranking
  \item Qualitative assessment
  \item Discontinued because:
    \begin{itemize}
    \item unable to turn this into competition
    \item still processing the data of the last two years
    \item restrictions of the BenchExec cluster (no disk space to store proofs)
    \end{itemize}
  \end{itemize}

\begin{frame}{Tracks, Solvers, Divisions, and Benchmarks}
  Teams: 20 (-5) %2023 25
  \bigskip

    \begin{tabular}{c|r@{}l|r@{}l|c}
    Track & \multicolumn{2}{c|}{Solvers} & \multicolumn{2}{c|}{Divisions}  & Benchmarks \\
    \hline
    Single Query  &  16&(-6)  & 19&(=)  & 123\,486 \\ %&(+10\,347)
    Incremental &  7&(=)   & 18&(+1)  & 22\,826   \\ %&(+525) Bitvec+Arith
    Unsat Core  &  6&(=)   & 19&(=)  & 66\,190  \\ %&(-6\,768)72\,958
    Model Validation  &  8&(-3)    &  7& (+ 8 exp.)  & 59\,625  \\ %-61\,083
%    Proof Exhibition  &  4&(=)    &  & 19 exp.  & 59\,114  \\
    \hline
    Parallel &   2&(-1)      &   &2 exp.  & 400 \\
    Cloud & 3&(+1)      &  &5 exp.  & 400 \\

@@

  Number in parenthesis shows changes from 2023
\end{frame}

@@

  SMT-COMP 2024 participants rely on multiple reasoning frameworks:
  \begin{itemize}
  \item CDCL(T), Saturation, MCSAT, CP
  \item automata
  \item finite domain
  \item local search
  \item besides wrappers extending the scope of existing solvers
  \end{itemize}

  \bigskip
  Six new solvers participated:
  \begin{itemize}
\item Algaroba {\footnotesize (Shah et al.)}
\item Amaya {\footnotesize (Havlena et al.)}
\item plat-smt {\footnotesize (Ewert)}
\item SMTS {\footnotesize (Kol\'{a}rik et al.)}
\item (STP-Parti-Bitwuzla {\footnotesize (Zhao et al.)})
\item Z3-Parti-Z3++ {\footnotesize (Zhao et al.)}
\end{itemize}

@@

\myincludepdf{amaya.pdf}

\myincludepdf{Bitwuzla.pdf}

\myincludepdf{cvc5.pdf}

\myvideopdf{OSTRICH}{Ostrich.mkv}

\myincludepdf{smtinterpol.pdf}

\myvideopdf{SMT-RAT}{SMT-RAT.mkv}

\myincludepdf{Yices2.pdf}

\myincludepdf{YicesQS.pdf}

\myincludepdf{z3-alpha.pdf}

\myincludepdf{Z3-Parti-Z3++.pdf}

\begin{frame}
    \frametitle{Other participants}
    \begin{itemize}
\item Algaroba
\vitem COLIBRI
\vitem iProver
\vitem OpenSMT
\vitem plat-smt
\vitem SMTS
\vitem STP
\vitem (STP-Parti-Bitwuzla)
\vitem Z3-Noodler
    \end{itemize}
  \end{frame}

@@

  Submitted by organisers
  \begin{itemize}
  \item Best solvers, per division, from previous years
  \end{itemize}
  \bigskip

  Submitted by participants
  \begin{itemize}
  \item Fixed solvers (Amaya, cvc5, OSTRICH, iProver)  
  \end{itemize}
  \bigskip
  \bigskip
  (All still in the process of running)
\end{frame}

@@

  \begin{itemize}
  \item  50411 benchmarks of 249\,301 have no status

  \vitem 2 benchmarks with disagreements (QF\_SLIA)

  \vitem We manually resolved the disagreements: authors confirmed solver
  unsoundness
  \vitem We had 4 solvers with soundness issues:
    \begin{itemize}
      \item Amaya (\& Fixed)
      \item cvc5 (\& Fixed)
      \item iProver (\& Fixed)
      \item OSTRICH (\& Fixed)
    \end{itemize}
  \end{itemize}
  \end{frame}

\begin{frame}[fragile]
  \frametitle{Plans for SMT-COMP 2025?}
  \begin{itemize}
    \item General: Documentation and streamlining
    \item General: More CI support for submissions
    \item Cloud/Parallel: Integrate them into the normal process
    \item MV: Algebraic number $\implies$ non experimental
\begin{verbatim}
(root-of-with-ordering (coeffs p_0 p_1 ... p_n) i)

@@

\begin{frame}{Plans for SMT-COMP 2025?}
  \begin{itemize}
    \item Proof \textit{validation} track
      \begin{itemize}
        \item We have to analyze the data still (830gb)
        \item Ideas to turn this into a competition are welcome
      \end{itemize}

@@

\begin{frame}{SMT-COMP organizing committee}
  Three people organize the SMT-COMP.  In 2024:
  \begin{itemize}
  \item Fran\c{c}ois Bobot
  \item Martin Bromberger
  \item Martin Jon\'{a}\v{s}
  \end{itemize}

  Martin B. has to leave early due to conflicts with his new job.\newline\newline
  Fran\c{c}ois will stay but only in an advisory role.
  \bigskip

  We need two successors for next year's competition.
  Contact us if you would like to volunteer!
\end{frame}

\begin{frame}{Acknowledgements}
  \begin{itemize}
    \item \emph{Andrea Micheli}: pysmt
    \item \emph{Guillaume Bury, FB.}: Model Validator
    \item \emph{Clark Barrett, Pascal Fontaine, Aina Niemetz, Mathias Preiner, Hans-Jörg Schurr}: SMT-LIB benchmarks
    \item \emph{Robert Jones, Benjamin F Jones, and team}: Cloud/Parallel Track
    \item \emph{Dirk Beyer and Philipp}: BenchExec and VerifierCloud support   
    \item \emph{Aaron Stump}: StarExec support
  \end{itemize}
  \bigskip

@@

\begin{frame}[shrink=0.95]{Benchmark contributors}
  In 2024 \emph{new benchmarks} were contributed by:

  \small
  \begin{itemize}
    \item TODO
  \end{itemize}
\end{frame}
